---
title: "Initial Korean covid 19"
author: "Jarrod Harris"
date: "Summer 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this 

```{r}
kr3 <- kRdata %>% group_by(sex,result) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
```

```{r}
basicC <- ggplot(kr3,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")
```





```{r}
basicC <- ggplot(kr3,aes(x=result,y=prcnt*100,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")
```

when you look at the  chart percentage males are dying more. At around 65% of people in the deceased column are males.
  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

when looking at the numbers 64 out of the 100% males are dead and 36 out of 100 females are dead. There are more males dead than females.



## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```




based off statistics of covid the males are dying more than the females percentage wise.

## Conclusions
By percent there were more females in the alive column than males. Also percentage wise males led in deaths. This shows that the death toll was not equivalent between males and females.  
